Mitigating Low Audio Quality by Domain Transfer for Better ASR

### 2020-04-15 2020-tsd-brno-denoise/denoise.tex
s/I/we/

\begin{abstract}
The paper presents an attempt to solve the problem of bad audio quality
affecting the accuracy of speech-to-text systems. I
describe an experiment with Gycle-GAN domain transfer from the defect to the
pure as a way to reduce the impact of noise, overdrive, reverberation and other
defects.
\keywords{speech recognition, GycleGAN, domain transfer, signal processing}
\end{abstract}

@@

I am looking into the problem of acoustic inconsistencies in terms of varying
overall quality, reverberation, noise, overdrive and other phenomena in a
dataset that is to be automatically transcribed and listened to by humans.

@@

I am conducting an experiment to make the input data similar to the training
data instead of the more usual other way around because 1) it is not easy to
make the volunteer annotators transcribe acoustically flawed data, 2) not only
speech recognition would benefit from successful denoising, the listening
experience would improve too, and 3) the model could stay leaner.

@@

There are two ways of dealing with this situation: 1) adapt CycleGAN so that it
can deal with a ``compact'' and a ``scattered'' domain or 2) cluster the
data and apply CycleGAN on the individual clusters. While adapting CycleGAN may be a
point of future work, I have explored the simpler way of clustering the data.

@@

In order to perform clustering on any dataset, a metric on the data is needed.
I have used that proposed by Mandel \& Ellis\cite{mandel2005song}, which
is based on mel-frequency cepstra. A distance matrix has been created using the
tool musly\cite{schnitzer2011using} and the clusters on top of that using
hierarchical clustering\cite{johnson1967hierarchical}.

@@

For curiosity, I looked at the distances of the files among each other and
compared it to that between other familiar sounds. For instance, two adjacent
segments of the same speaker from a meeting of the Czech parliament has a
distance of 1.4. Different speakers have a distance of 6.5. One of these
segments compared to a black metal song chorus has the distance of 519. On the
other hand, the median distance among the Spoken corpus of Karel Makoň, where
there is a single speaker, amounts to 55.9.

@@

I have used the voice transfer method as proposed by Kaneko \&
Kameoka\cite{kaneko2017parallel} and implemented by
Mao\footnote{github.com/leimao/Voice\_Converter\_CycleGAN}. I have trained
transfer 1) between the clean and the overdriven cluster and 2) between the
clean and the low-speed cluster.

After 200 epochs, I have performed speech recognition on the transferred data.
The result is summed up in Table~\ref{tab:results}.

@@

The small but apparent reduction in word error rate in case of the overdriven
cluster is the positive, promising result. The improvement is rather small, and
as such it does not seem worthwhile applying it on the entire corpus. It
proves, however, that there is some potential. An important immediate benefit is
that the transferred overdriven (or should I say de-overdriven) sound files are
much easier on the ear and it can be a help for the human listener.

>>>>

\begin{abstract}
The paper presents an attempt to solve the problem of bad audio quality
affecting the accuracy of speech-to-text systems. We
describe an experiment with Gycle-GAN domain transfer from the defect to the
pure as a way to reduce the impact of noise, overdrive, reverberation and other
defects.
\keywords{speech recognition, GycleGAN, domain transfer, signal processing}
\end{abstract}

@@

We are looking into the problem of acoustic inconsistencies in terms of varying
overall quality, reverberation, noise, overdrive and other phenomena in a
dataset that is to be automatically transcribed and listened to by humans.

@@

We are conducting an experiment to make the input data similar to the training
data instead of the more usual other way around because 1) it is not easy to
make the volunteer annotators transcribe acoustically flawed data, 2) not only
speech recognition would benefit from successful denoising, the listening
experience would improve too, and 3) the model could stay leaner.

@@

There are two ways of dealing with this situation: 1) adapt CycleGAN so that it
can deal with a ``compact'' and a ``scattered'' domain or 2) cluster the
data and apply CycleGAN on the individual clusters. While adapting CycleGAN may be a
point of future work, we have explored the simpler way of clustering the data.

@@

In order to perform clustering on any dataset, a metric on the data is needed.
We have used that proposed by Mandel \& Ellis\cite{mandel2005song}, which
is based on mel-frequency cepstra. A distance matrix has been created using the
tool musly\cite{schnitzer2011using} and the clusters on top of that using
hierarchical clustering\cite{johnson1967hierarchical}.

@@

For curiosity, we looked at the distances of the files among each other and
compared it to that between other familiar sounds. For instance, two adjacent
segments of the same speaker from a meeting of the Czech parliament has a
distance of 1.4. Different speakers have a distance of 6.5. One of these
segments compared to a black metal song chorus has the distance of 519. On the
other hand, the median distance among the Spoken corpus of Karel Makoň, where
there is a single speaker, amounts to 55.9.

@@

We have used the voice transfer method as proposed by Kaneko \&
Kameoka\cite{kaneko2017parallel} and implemented by
Mao\footnote{github.com/leimao/Voice\_Converter\_CycleGAN}. We have trained
transfer 1) between the clean and the overdriven cluster and 2) between the
clean and the low-speed cluster.

After 200 epochs, we have performed speech recognition on the transferred data.
The result is summed up in Table~\ref{tab:results}.

@@

The small but apparent reduction in word error rate in case of the overdriven
cluster is the positive, promising result. The improvement is rather small, and
as such it does not seem worthwhile applying it on the entire corpus. It
proves, however, that there is some potential. An important immediate benefit is
that the transferred overdriven (or should we say de-overdriven) sound files are
much easier on the ear and it can be a help for human listeners.
